Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis

### 2024-04-03 paper/paper.qmd
upload

The remainder of this paper is structured as follows. @sec-data introduces the data used for analysis and findings, including visualizations of the variables of interest, @sec-model
proposes a straightforward linear regression model to examine and forecasts the effects of time (year) and geographical location (province) on income inequality in Canada.In @sec-results, we display the interpretations of the models alongside other findings from analyzing the data. @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.

@@

Overall, our estimand focuses on uncovering the nuanced relationship between time, geographical location, and income inequality in Canada, providing insights into the factors influencing income distribution dynamics over the study period.

@@

Upper income limit, income share and average income by economic family type and income decile

@@

The estimates for Upper Income Limit, After-Tax Income by Economic and Income Decile in Canada are based on data from various surveys spanning different periods. From 1976 to 1992, data was collected from the Survey of Consumer Finances (SCF). For the period from 1993 to 1997, a combination of SCF and the Survey of Labour and Income Dynamics (SLID) was used. Subsequently, from 1998 to 2011, data solely from SLID was utilized. Starting from 2012, the Canadian Income Survey (CIS) became the primary data source. More details on these surveys and their revisions can be found in publications by Statistics Canada, including "Revisions to 2006 to 2011 Income Data" (2015) by Statistics Canada and other related papers by Cotton (2000) and Lathe (2005).

@@

6.  **Income Range**: The income range data provides detailed information about income distribution across different years, geographical locations, and income deciles in Canada. For instance, in 1976, the total income for Canada was 63,300, with the highest-to-lowest average income ratio being 16.8, indicating a significant income gap between the highest and lowest deciles. The income range for the lowest decile in Canada was 9,400, while the income for the highest decile was $157,500.

@@

- $\beta_{2}$ is the coefficient associated with the "Geographical Location" variable (\(Province_{i}\)), representing the average difference in income range between different provinces, holding the year constant.

>>>>

The remainder of this paper is structured as follows. @sec-data introduces the data used for analysis and findings, including visualizations of the variables of interest, @sec-model
proposes a straightforward linear regression model to examine and forecasts the effects of time (year) and geographical location (province) on income inequality in Canada. In @sec-results, we display the interpretations of the models alongside other findings from analyzing the data. @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.

@@

Overall, our estimand focuses on uncovering the nuanced relationship between time, geographical location, and income inequality in Canada.

@@

According to a report by  @statscanada, titled \textit{Upper income limit, income share and average income by economic family type and income decile},the data provided valuable insights into income distribution patterns in Canada.

@@

The estimates for \textit{Upper income limit, income share and average income by economic family type and income decile} are based on data from various surveys spanning different periods. From 1976 to 1992, data was collected from the Survey of Consumer Finances (SCF). For the period from 1993 to 1997, a combination of SCF and the Survey of Labour and Income Dynamics (SLID) was used. Subsequently, from 1998 to 2011, data solely from SLID was utilized. Starting from 2012, the Canadian Income Survey (CIS) became the primary data source. More details on these surveys and their revisions can be found in publications by Statistics Canada, including "Revisions to 2006 to 2011 Income Data" (2015) by Statistics Canada and other related papers by Cotton (2000) and Lathe (2005).

@@

6.  **Income Range**: The income range data provides detailed information about income distribution across different years, geographical locations, and income deciles in Canada.

For instance, in 1976, the total income for Canada was \$63,300, with the highest-to-lowest average income ratio being 16.8, indicating a significant income gap between the highest and lowest deciles. The average after-tax income for the lowest decile in Canada was \$9,400, while the income for the highest decile was \$157,500.

@@

- $\beta_{2}$ is the coefficient associated with the "Geographical Location" variable ($Province_{i}$), representing the average difference in income range between different provinces, holding the year constant.
